---
title: "BCP MLU Preliminary Analysis"
author: "Trevor KM Day"
date: "June 25, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

Packages: `tidyverse, lme4, MuMIn, viridis, gridExtra`.

```{r load-libraries, include=FALSE}

if(.Platform$OS.type == "unix") {
  setwd("/Volumes/GoogleDrive/My Drive/Research/MCDI/MCDI-analysis")
} else {
  setwd("G:/My Drive/Research/MCDI/MCDI-analysis/")
}

# Libraries

library(tidyverse)
library(lme4)
library(MuMIn)
library(viridis)
library(gridExtra)

# Helper functions

# source("wordbank-functions.R")
# source("format-BCP-funcs.R")

date <- "200609"

```

# Load data

Wordbank had only age, sex, and mother's education as demographics. Here, we 
show those for the available subjects, as well as the correlation between MLU3m
and MLU3w.

```{r load-data, echo=FALSE, message=FALSE, warning=FALSE}

# Load data

#BCP_WG_scored <- readRDS("data/BCP_WG_scored.rds")
BCP_WS_scored <- readRDS(paste0("data/BCP_WS_scored-", date, ".rds"))[[1]] %>%
                  select(-how_use_words) %>%
                  mutate(LEX = action_words + animals + body_parts + clothing +
                                descriptive_words + food_drink +
                                furniture_rooms + games_routines +
                                helping_verbs + household + outside + people +
                                places + toys + vehicles + sounds,
                         SYN = pronouns + quantifiers + question_words +
                                time_words + word_endings_nouns +
                                word_endings_verbs + word_forms_nouns +
                                word_forms_verbs + complexity +
                                connecting_words + locations,
                         partI = action_words + animals + body_parts + clothing +
                           descriptive_words + food_drink +
                           furniture_rooms + games_routines +
                           helping_verbs + household + outside + people +
                           places + toys + vehicles + sounds + pronouns +
                           quantifiers + question_words + time_words +
                           connecting_words + locations,
                         partII = word_endings_nouns + word_endings_verbs +
                           word_forms_nouns + word_forms_verbs + complexity)

# Read demographics file 
# Explicit levels in momed set order
BCP_demo <- read_csv(paste0("data/BCP-demographics-", date, ".csv")) %>%
              select(CandID, sex, educ_momed, educ_momed_n, income_inr,
                     family_size) %>%
              mutate(educ_momed = replace(educ_momed,
                                          educ_momed == "not_answered",
                                          NA) %>%
                                  factor(levels = c("some_high", "high", 
                                                    "some_college", "college",
                                                    "some_grad", "grad",
                                                    "not_answered"))) 

# Fill in missing INR based on mother's education and family size
inr.predictor <- lm(income_inr ~ 1 + educ_momed_n + family_size, 
                    data = filter(BCP_demo, !is.na(income_inr)))
inr.predicted <- predict(inr.predictor, 
                         newdata = filter(BCP_demo, is.na(income_inr)))

# There are some not calculated - but they don't have MLU. If needed, could be
# replaced with mean/median INR
BCP_demo$income_inr[is.na(BCP_demo$income_inr)] <- inr.predicted

# MLU file
BCP_MLU <- readRDS(paste0("data/BCP_WS_MLU3-", date, ".rds"))

# Join them all into "with MLU file"
wMLU <- right_join(BCP_WS_scored, BCP_MLU) %>%
          left_join(., BCP_demo, by = c("data_id" = "CandID"))


p1 <- ggplot(wMLU, aes(x = MLU3w, y = MLU3m, color = age)) +
        geom_point() +
        geom_line(aes(group = data_id)) +
        scale_color_viridis() + 
        labs(x = "Words", y = "Morphemes") +
        theme(legend.position = "bottom")

p2 <- ggplot(BCP_demo, aes(educ_momed)) +
        geom_histogram(stat = "count") +
        scale_x_discrete(labels = c("Some Sec. (10)", "Secondary (12)", 
                                    "Some College (14)", "College (16)", 
                                    "Some Grad (18)", "Grad (20)", 
                                    "Not Answered")) +
        labs(x = "Level", y = "Count") + 
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

p3 <- ggplot(wMLU, aes(x = age)) +
        geom_histogram(binwidth = 1) +
        scale_x_continuous(breaks = seq(19, 38, by = 2)) +
        labs(x = "Age", y = "Count")


p4 <- ggplot(wMLU, aes(x = sex, fill = sex)) +
        geom_histogram(stat = "count") +
        labs(x = "Gender", y = "Count") +
        theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4,
             layout_matrix = rbind(c(2,  3, 1),
                                   c(NA, 4, 1)))


# Clean up
rm(BCP_demo, BCP_MLU, BCP_WS_scored, p1, p2, p3, p4)


means <- wMLU %>%
          select(age, educ_momed_n, LEX, SYN, partI, partII, MLU3m, MLU3w) %>%
          summarize_all(list(M = mean))

sds <- wMLU %>%
          select(age, educ_momed_n, LEX, SYN, partI, partII, MLU3m, MLU3w) %>%
          summarize_all(list(SD = sd))

var.sum <- rbind(unname(unlist(means)), unname(unlist(sds))) %>%
            round(1)
colnames(var.sum) <- c("Age", "Mom Ed", "LEX", "SYN", "Part I", "PartII",
                       "MLU3m", "MLU3w")
rownames(var.sum) <- c("Mean", "SD")

var.sum

```

# Analysis

Since MLU3m and MLU3w are so closely correlated, I use MLU3m for here on out.

```{r include=FALSE}

model.lex_m <- lmer(MLU3m ~ LEX + (1|data_id), data = wMLU, REML = TRUE)
model.syn_m = lmer(MLU3m ~ SYN + (1|data_id), data = wMLU, REML = TRUE)

# Control model

model.lex_m_ctrl <- lmer(MLU3m ~ LEX + age + sex + educ_momed + (1|data_id),
                          data = wMLU, REML = TRUE)
model.syn_m_ctrl <- lmer(MLU3m ~ SYN + age + sex + educ_momed + (1|data_id),
                          data = wMLU, REML = TRUE)

#
# Compare with MCDI categories
#

# mcdi categories are word_* + complexity vs. all others

model.partI_m <- lmer(MLU3m ~ partI + (1|data_id), data = wMLU, REML = TRUE)
model.partII_m <- lmer(MLU3m ~ partII + (1|data_id), data = wMLU, REML = TRUE)

model.partI_m_ctrl <- lmer(MLU3m ~ partI + age + sex + educ_momed + (1|data_id),
                            data = wMLU, REML = TRUE)
model.partII_m_ctrl <- lmer(MLU3m ~ partII + age + sex + educ_momed +(1|data_id),
                              data = wMLU, REML = TRUE)

AICc_noctrl <- sapply(list(model.lex_m, model.syn_m, model.partI_m,
                           model.partII_m),
                      AICc) %>%
                round()

AICc_ctrl <- sapply(list(model.lex_m_ctrl, model.syn_m_ctrl, model.partI_m_ctrl,
                           model.partII_m_ctrl),
                      AICc) %>%
                round()

AICc_tbl <- rbind(AICc_noctrl, AICc_ctrl)
colnames(AICc_tbl) <- c("LEX", "SYN", "Part I", "Part II")

```

So we see that control overall improves the models, and that SYN/Part II are 
better than LEX/Part I, but SYN doesn't meaningfully improve over Part II.

# Reanalysis

However, it turns out SYN is a superset of Part II, so we can do a direct ANOVA
for more information.

```{r}

# Demographic predictors
demo.predictors <- c("age", "sex", "educ_momed", "income_inr")

# All part I/lexical predictors
lexical.predictors <- c("action_words", "animals", "body_parts", "clothing",
                        "descriptive_words", "food_drink", "furniture_rooms", 
                        "games_routines", "helping_verbs", "household", 
                        "outside", "people", "places", "toys", "vehicles", 
                        "sounds")

# Predictors in Part II, but also syntax
swap.predictors <- c("pronouns ", "quantifiers", "question_words", "time_words", 
                      "word_endings_nouns", "connecting_words", "locations")

# Part II/syntax
partII.predictors <- c("word_endings_verbs", "word_forms_nouns",
                        "word_forms_verbs", "complexity")

formula.syntax_demo <- paste("MLU3m ~ ", 
                             paste(c(demo.predictors, "(1|data_id)"),
                                    collapse = " + "))

formula.syntax_partII <- paste("MLU3m ~ ", 
                                paste(c(demo.predictors, partII.predictors),
                                      collapse = " + "),
                                "+ (1|data_id)")

formula.syntax_SYN <- paste("MLU3m ~", 
                            paste(c(demo.predictors, swap.predictors, 
                                    partII.predictors), 
                                  collapse = " + "),
                            "+ (1|data_id)")

syntax00 <- lmer(MLU3m ~ 1 + (1|data_id), data = wMLU, REML = TRUE)
syntax0  <- lmer(formula = formula.syntax_demo, data = wMLU, REML = TRUE)
syntax1  <- lmer(formula = formula.syntax_partII, data = wMLU, REML = TRUE)
syntax2  <- lmer(formula = formula.syntax_SYN, data = wMLU, REML = TRUE)

anova(syntax00, syntax0, syntax1, syntax2)

# Improvements:
# demo   >  intercept-only
# partII >  demo
# SYN    !> partII

```

# Conclusion

Preliminary analysis with `r nrow(wMLU)` subjects shows that MLU3m is 
differentially associated with syntax metrics over lexical metrics. However,
the proposed metric is not significantly better than traditional MCDI Part II.

Analysis plan:

 - Identify other variables and test them further against BCP82 sample.
 - Once phenoscreening is done, use any variables identified above and rerun.

\newpage
# Package info

```{r package-info}

version

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]

```