---
title: "Growth Curving"
author: "Trevor K.M. Day"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
bibliography: growth-curving-bib.json
---


```{r setup, include=FALSE}

if (.Platform$OS.type == "unix") {
  setwd("/Volumes/GoogleDrive/My Drive/Research/MCDI/MCDI-analysis/code/growth_curving")
} else {
  setwd("G:/My Drive/Research/MCDI/MCDI-analysis/code/growth_curving")
}

knitr::opts_chunk$set(echo = TRUE)
```

# Setup

Load packages `tidyverse`, `plyr`, `ggplot2`, `scales`, `ggrepel`, `gridExtra`,
`growthcurver`, `educate` (from Andy).

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse) 
library(plyr)
library(ggplot2)
library(scales)

library(ggrepel)
library(gridExtra)

library(growthcurver)

# library(devtools)
# devtools::install_github("zief0002/educate")
library(educate)

select <- dplyr::select

```

Load the first person with five points

```{r data}

test_subjs <- "G:/My Drive/Research/MCDI/MCDI-analysis/data/other/test-subjs.csv"

test <- read_csv(test_subjs) %>%
          arrange(data_id, age) %>%
          filter(data_id == 176427)

test_ns <- read_csv(test_subjs) %>%
  group_by(data_id) %>%
  nest() %>%
  mutate(
    nobs = map(data, nrow)
  )

```

# Introduction

Throughout, I will be using notation from @tjorve_and_tjorve, where:

 - $W_0$: The starting population. For our purposes, can be estimated at 1 or 
          less.
 - $A$: The upper asymptote, here I use $680+1$, the size of MCDI plus one so
          the participant is actually modeled as learning all words.
 - $t$: Time. Here, age in months.
 - $k_g$: *Relative* (or *intrinsic*) growth rate.
 - $k_U$: Absolute maximum growth rate (at inflection). 
          Given as $k_U=\frac{A\times{}k_g}{e}$. 
          Has units $\frac{\textrm{words}}{\textrm{month}}$.
 - $W_i,T_i$: Value and time at inflection. $W_i$ is fixed at $\frac{A}{e}\approx251$,
                and $T_i$ can be solved for. 
                
```{r values}

# Upper asymptote
A <- 681

# Lower asymptote: the smallest value R can represent
W_0 <- .Machine$double.eps

```
                
# Basic Gompertz

The Gompertz equation given by `growthcurver`, converted to the new notation is: 

<!-- $$ -->
<!-- y = \frac{K}{1 + \frac{K - N_0}{N_0}e^{-rx}} -->
<!-- $$ -->

$$
y = \frac{A}{1 + \frac{A - W_0}{W_0}(\exp{(-k_g\times{}x)})}
$$



This can be created like so:

```{r growthcurver}
gomp.fit <- SummarizeGrowth(test$age, test$inventory)
gomp.fit
```

We write a quick function to calculate the fit over a given range. We use 
`0.01` as a step as a day is `0.03` months. See the first part of 
`predict.gomp`.

```{r gomp_fit}

predict.gomp <- function(fit = NULL, range = seq(0, 36, by = 0.01), max = 681,
                         ci = FALSE, kg = NULL, kg.se = NULL) {
  
  if (class(fit) == "gcfit") {
    
    # Result from growth curver
    A <- fit$vals$k
    W_0 <- fit$vals$n0
    k_g <- fit$vals$r
    k_g.se <- fit$vals$r_se
    
    if (ci)
      k_g.se <- 1.96 * k_g.se

    result <- A/(1 + ((A - W_0)/W_0) * exp(-1 * k_g * range))
    result.lo <- A/(1 + ((A - W_0)/W_0) * exp(-1 * (k_g - k_g.se) * range))
    result.hi <- A/(1 + ((A - W_0)/W_0) * exp(-1 * (k_g + k_g.se) * range))
    
  } else if (class(fit) == "nls") {
    
    # Result from my constrained model
    A <- max
    W_0 <- .Machine$double.eps
    k_g <- coef(fit) %>% unname()
    k_g.se <- summary(fit)$coefficients[1, 2]
    
    if (ci)
      k_g.se <- 1.96 * k_g.se

    result <- W_0 * (A / W_0) ^ (1 - exp(-1 * k_g * range))
    result.lo <- W_0 * (A / W_0) ^ (1 - exp(-1 * (k_g - k_g.se) * range))
    result.hi <- W_0 * (A / W_0) ^ (1 - exp(-1 * (k_g + k_g.se) * range))
    
  } else if (class(fit) == "NULL") {
    
    if (!is.null(kg) & !is.null(kg.se)) {
      
      A <- max
      W_0 <- .Machine$double.eps
      k_g <- kg
      k_g.se <- kg.se 
      
      if (ci)
        k_g.se <- 1.96 * k_g.se
    

      result <- W_0 * (A / W_0) ^ (1 - exp(-1 * k_g * range))
      result.lo <- W_0 * (A / W_0) ^ (1 - exp(-1 * (k_g - k_g.se) * range))
      result.hi <- W_0 * (A / W_0) ^ (1 - exp(-1 * (k_g + k_g.se) * range))
      
      
    }
    
  }

  
  out <- cbind(range, result, result.lo, result.hi) %>%
          as.data.frame()
  
  return(out)
  
}

```

```{r plot_1, fig.height=3, fig.width=5, fig.cap="Gompertz curve from growthcurver package", echo=FALSE}

fit1 <- predict.gomp(gomp.fit)

ggplot(NULL) +
  scale_x_continuous(limits = c(10, 36), breaks = seq(10, 36, by = 2)) +
  scale_y_continuous(limits = c(0, 680)) +
  geom_ribbon(data = fit1, aes(x = range, ymin = result.lo, ymax = result.hi),
              fill = "lightblue", alpha = 0.5) +
  geom_line(data = fit1, aes(x = range, y = result), color = "blue") +
  geom_point(data = test, aes(x = age, y = inventory)) +
  labs(x = "Age (mo.)", y = "Inventory") +
  geom_hline(yintercept = 680)

```

The standard error of $k_g$ is used for two additional lines and plotted here as 
an error range.

However, this has the limitation that $A$ is 
estimated by the function, which 
means it may pick an asymptote well below 680.

I also want to point out this is a remarkably good fit; other subjects are not
so good.

# Gompertz II: Electric Pertzaloo

Through trial and error, I identified a form of the Gompertz equation that 
works well with `nls` and allows $A$ and $W_0$ as parameters. As a result, only
$k_g$ has to converge for this model.

For $A$, we use $680+1$, as otherwie, a kid who never learns the 680th word
on the MCDI would be modeled. Although, as a reminder, we are most interested
in the shape of the first two-thirds of the model, since the greater asymptote
is an artifact of the instrument.

This equation is:

$$
y = W_0\left(\frac{A}{W_0}\right)^{1-\exp(-k_G \times{} t)}
$$

The absolute growth rate, $k_u$, which technically has the units words/month,
can be solved for as follows:

$$
k_u = \frac{A\times{}k_g}{e} = \frac{681}{e}k_g \approx 251\times{}k_g
$$

Function and solver below:

```{r gomp2}
gomp2.fit <- function(data, response = "inventory", max = 681, max.iter = 50) {
  
  A <- max
  W_0 <- .Machine$double.eps
  
  # This is the formula that works best to solve in R
  # 19 refers to its number in the Tjorve and Tjorve paper
  fit19 <- NULL
  try( fit19 <- nls(as.formula(paste(response, 
                                "~ W_0 * (A / W_0) ^ (1 - exp(-k_g * age))")),
                   data = data,
                   start = list(k_g = .1),
                   control = list(maxiter = max.iter)) )
  
  if (is.null(fit19))
    fit19 <- NA
  
  return(fit19)
  
}

# Estimate relative growth rate
gomp2 <- gomp2.fit(test)
gomp2.kg <- summary(gomp2)$coefficients[1, 1]

# Solve for abs growth rate
gomp2.kU <- 681 * gomp2.kg / exp(1)

# Solve for a given X or Y, given kg
solve.gomp2 <- function(x = NA, y = NA, k_g, A = 681) {
  
  W_0 <- .Machine$double.eps
  
  if (is.na(y) & !is.na(x))
    result <-  W_0 * (A / W_0) ^ (1 - exp(-k_g * x))
  else if (is.na(x) & !is.na(y))
    result <- -log(1 - log(y / W_0) / log(A / W_0)) / k_g
  else
    message("Exactly one of x/y must be specified")
  
  return(result)
  
}

```

Now we plot the Gompertz fit, including the point of inflection, and the 
maximum absolute growth rate, the rate at $\{T_i, W_i\}$. The value of $W_i$,
$\frac{A}{e}=251$ is a feature of the model, so we solve for $T_i$ given a 
$k_g$. $k_u$ is the slope of the tangent line at that point.


```{r plot2, echo=FALSE, fig.height=4, fig.width=6}

# Value at inflection: constant (a feature of the Gompertz formula)
W_i <- 681 / exp(1)

# Time at inflection
# Slope at inflection is gomp2.kU
T_i <- solve.gomp2(y = W_i, k_g = gomp2.kg)

# Solve for intercept of tangent line
b <- W_i - gomp2.kU * T_i

# This block solves for upper and lower estimates 
# gomp2.kg.se <- summary(gomp2)$coefficients[1, 2]
# gomp2.kgs <- c(gomp2.kg - gomp2.kg.se, gomp2.kg + gomp2.kg.se)

colors <- hue_pal()(3)

ggplot(NULL) +
  geom_point(data = test, 
             aes(x = age, y = inventory),
             alpha = 0.75,
             color = colors[1]) +
  geom_line(data = test, 
             aes(x = age, y = inventory),
             alpha = 0.75,
             color = colors[1],
             size = 1) +
  scale_x_continuous(limits = c(0, 60), breaks = seq(10, 36, by = 4), 
                     minor_breaks = NULL) +
  scale_y_continuous(limits = c(NA, 700)) + 
  labs(x = "Age (mo.)", y = "Inventory") +
  geom_hline(yintercept = c(680, 680 / exp(1)), size = 1, linetype = "dashed", 
             color = "darkred") +
  stat_function(data = test,
                fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-gomp2.kg * age)) },
                color = colors[2],
                size = 1.1,
                alpha = 0.75) +
  geom_abline(intercept = b,
              slope = gomp2.kU,
              color = "black",
              size = 0.75,
              alpha = 0.5) +
  geom_point(aes(x = T_i, y = W_i), 
             size = 3, 
             shape = 16, 
             color = colors[3]) +
  geom_label_repel(aes(x = T_i, y = W_i), 
                    size = 4, 
                    label = paste0(round(T_i), ",", round(W_i)),
                    nudge_x = 5,
                    color = colors[3]) +
  annotate("text",
           x = 15, y = 600, 
           label = paste0("k_U=", round(gomp2.kU, 1))) +
  theme_bw()
  

```

```{r iterate-for-pres, include=FALSE}

# This is an animation for a presentation.

p1 <- ggplot(NULL) +
        geom_point(data = test, 
                   aes(x = age, y = inventory),
                   size = 2,
                   color = "black") +
        geom_line(data = test, 
                   aes(x = age, y = inventory),
                   color = "black",
                   size = 1) +
        scale_x_continuous(limits = c(0, 60), 
                           breaks = seq(10, 36, by = 2), 
                           minor_breaks = NULL) +
        scale_y_continuous(limits = c(NA, 700)) + 
        labs(x = "Age (mo.)", y = "Inventory") +
        geom_hline(yintercept = 680, size = 1, linetype = "dashed", 
                   color = "darkred")

p2 <- p1 +
        stat_function(fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-gomp2.kg * age)) },
                      color = colors[2],
                      size = 1.1,
                      alpha = 0.75)

p3 <- p2 +
        geom_hline(yintercept = 680 / exp(1), size = 1, linetype = "dashed", 
                    color = "darkred")

p4 <- p3 +  
        geom_abline(intercept = b,
                    slope = gomp2.kU,
                    color = "red",
                    size = 1) +
        geom_label_repel(aes(x = T_i, y = W_i), 
                          size = 4, 
                          label = paste0(round(T_i), ",", round(W_i)),
                          nudge_x = 5,
                          color = "red") +
        annotate("text",
                 x = 15, y = 600, 
                 label = paste0("k_U=", round(gomp2.kU, 1)))


dir.create("plots/", showWarnings = TRUE)
names <- paste0("plots/gompertz", 1:4, ".png")
plots <- list(p1, p2, p3, p4)

for (i in 1:4) {
  
  png(names[i], width = 6, height = 5, res = 300, units = "in")
  print(plots[[i]])
  dev.off()
  
}

```

## Reducing observations

Here we are modeling five points. However, not everyone has five points, so 
how few can we use?

```{r five_or_less, echo = FALSE}

gomp2s <- lapply(5:3, function(x) gomp2.fit(data = test[1:x, ]))
gomp2.kgs <- sapply(gomp2s, function(x) summary(x)$coefficients[1, 1])
gomp2.kUs <- A * gomp2.kgs / exp(1)

T_is <- sapply(gomp2.kgs, function(x) solve.gomp2(y = W_i, k_g = x))
bs <- W_i - gomp2.kUs * T_is

#colors <- hue_pal()(3)

ggplot(NULL) +
  geom_point(data = test, 
             aes(x = age, y = inventory),
             alpha = 0.75,
             color = "black") +
  geom_line(data = test, 
             aes(x = age, y = inventory),
             alpha = 0.75,
             color = "black",
             size = 1) +
  scale_x_continuous(limits = c(0, 60), breaks = seq(10, 36, by = 2), 
                     minor_breaks = NULL) +
  scale_y_continuous(limits = c(NA, 700)) + 
  labs(x = "Age (mo.)", y = "Inventory") +
  geom_hline(yintercept = c(680, 680 / exp(1)), size = 1, linetype = "dashed", 
             color = "darkred") +
  stat_function(data = test, 
                fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-gomp2.kgs[1] * age)) },
                color = colors[1],
                size = 1.1,
                alpha = 0.5) +
  stat_function(data = test, 
                fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-gomp2.kgs[2] * age)) },
                color = colors[2],
                size = 1.1,
                alpha = 0.5) +
  stat_function(data = test, 
                fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-gomp2.kgs[3] * age)) },
                color = colors[3],
                size = 1.1,
                alpha = 0.5) +
  geom_abline(intercept = bs,
              slope = gomp2.kUs,
              color = colors,
              size = 0.75,
              alpha = 0.5) +
  geom_point(aes(x = T_is, y = W_i), 
             size = 3, 
             shape = 16,
             color = colors) +
  geom_label_repel(aes(x = T_is, y = W_i), 
                    size = 4, 
                    label = paste0(round(T_is, 1), ",", round(W_i)),
                    nudge_x = 5,
                   force = 25, 
                    color = colors) +
  annotate("text",
           x = 15, y = 600 + c(0, 25, 50), 
           label = paste0("k_U=", round(gomp2.kUs, 2)),
           color = colors) +
  theme_bw()

```

It looks like four and five give reasonably close values, with three not doing
too poorly for this individual.

## Modeling across subjects

```{r load-all-data}

# Load data as a list
test_all <- read_csv(test_subjs) %>%
              arrange(data_id, age) %>%
              split(., f = .$data_id)

```

The `test` file contains data from `r length(test_all)` subjects who have at 
least four 
observations. Here, we're going to repeat the same steps as above and plot all
51 to examine how consistent curves are with other data.


I'm going to invisibly fit the curves here, with tangent lines and $\{T_i,W_i\}$
coordinates on the left and SE envelopes on the right.

```{r fit-curves, echo=FALSE, fig.height=3.5, fig.width=6.5, message=FALSE, warning=FALSE}

all.fits5 <- lapply(test_all, function(x) if (nrow(x) == 5) { gomp2.fit(x) } else { NA })
all.fits4 <- lapply(test_all, function(x) gomp2.fit(x[1:4, ]))
all.fits3 <- lapply(test_all, function(x) gomp2.fit(x[1:3, ]))

# Plots three curves/tangents, list fits 3 -> 5
plot.tri <- function(data, fit3, fit4, fit5 = NA) {
  
  if (is.list(fit5)) {
    fits <- list(fit3, fit4, fit5)
  } else {
    fits <- list(fit3, fit4)
  }
  
  kgs <- sapply(fits, function(x) summary(x)$coefficients[1, 1])
  
  # The maximum number of colors is 3
  colors <- hue_pal()(3)
  
  p <- ggplot(NULL) +
        geom_point(data = data, 
                   aes(x = age, y = inventory),
                   alpha = 0.75,
                   color = "black") +
        geom_line(data = data, 
                   aes(x = age, y = inventory),
                   alpha = 0.75,
                   color = "black",
                   size = 1) +
        scale_x_continuous(limits = c(0, 60), 
                           breaks = seq(10, 36, by = 4), 
                           minor_breaks = NULL) +
        scale_y_continuous(limits = c(NA, 700)) + 
        labs(x = "Age (mo.)", y = "Inventory") +
        geom_hline(yintercept = c(680, 680 / exp(1)), 
                   size = 1, 
                   linetype = "dashed", 
                   color = "darkred") 
  
  # Here are the basic curves
  curves <- lapply(1:length(fits), function(x)
                      stat_function(data = data,
                                    fun = function(age) { W_0 * (A / W_0) ^ (1 - exp(-kgs[x] * age)) },
                                    color = colors[x],
                                    size = 1.1,
                                    alpha = 0.5))
  
  # Add the basic curves to the plot (only add 3 if it exists)
  p2 <- p + curves[[1]] + curves[[2]]
  if (length(curves) == 3) {
    p2 <- p2 + curves[[3]]
  }
  
  # Ranges using predict.gomp
  ribbons <- lapply(fits, predict.gomp)
  
  geom_ribbons <- lapply(1:length(ribbons), 
                         function(x) {
    
                            dat <- ribbons[[x]]
      
                            r <- geom_ribbon(data = dat, 
                                              aes(x = range, 
                                                  ymin = result.lo, 
                                                  ymax = result.hi), 
                                                  fill = colors[x],
                                                  alpha = 0.5)
                            
                            return(r)
                          
                          } )
  
  
  kUs <- 681 * kgs / exp(1)
  T_is <- sapply(kgs, function(x) solve.gomp2(y = W_i, k_g = x))
  bs <- W_i - kUs * T_is
  
  tangents <- list(geom_abline(intercept = bs,
                    slope = kUs,
                    color = colors[1:length(bs)],
                    size = 0.75,
                    alpha = 0.5),
                  geom_point(aes(x = T_is, y = W_i),
                             size = 3,
                             shape = 16,
                             color = colors[1:length(T_is)]),
                  geom_label_repel(aes(x = T_is, y = W_i),
                                    size = 4,
                                    label = paste0(round(T_is, 1), ",", round(W_i)),
                                    nudge_x = 5,
                                    force = 25,
                                    color = colors[1:length(T_is)]) )
  
  p3 <- p2 + tangents
  
  y_pos <- (600 + c(0, 25, 50))[1:length(kUs)]
  labels <- paste0(2 + 1:length(kUs), ": k_U=", round(kUs, 2)) 
  
  annot <- list(annotate("text",
                           x = 15, y = y_pos,
                           label = labels,
                           color = colors[1:length(kUs)]),
                  theme_bw())
  
  p4 <- p3 + annot
  pr1 <- p + geom_ribbons + curves + annot

  grid.arrange(p4, pr1, ncol = 2, top = unique(data$data_id))
  
}

# plot.tri(test_all[[51]], 
#          fit3 = all.fits3[[51]], 
#          fit4 = all.fits4[[51]], 
#          fit5 = all.fits5[[51]])

all.plots <- lapply(1:51, function(x) plot.tri(test_all[[x]],
                                                all.fits3[[x]],
                                                all.fits4[[x]],
                                                all.fits5[[x]]))

for (p in all.plots){
  p
}

```

## Distribution of $k_u$

How widely distributed is $k_u$? In other words, are we getting meaningful
variation in $k_u$, or does everyone develop at the same rate?
Boxplots are shown to visualize variation in estimates per-person, per number of
points.

```{r dist-ku}

kgs5 <- sapply(all.fits5, function(x) if (is.list(x)) {summary(x)$coefficients[1, 1]} else {NA})
kgs4 <- sapply(all.fits4, function(x) summary(x)$coefficients[1, 1])
kgs3 <- sapply(all.fits3, function(x) summary(x)$coefficients[1, 1])

all.kgs <- cbind(kgs3, kgs4, kgs5)
all.kUs <- (A * all.kgs / exp(1)) %>%
            as.data.frame() %>% 
            rownames_to_column() %>%
            pivot_longer(-rowname) %>%
            na.omit()

all.kUs %>%
  filter(
    name == "kgs4"
  ) %>%
  ggplot(aes(value)) +
    geom_density()

# max_vals <- bind_rows(test_all) %>%
#               group_by(data_id) %>%
#               summarize(age = max(age), W = max(inventory)) %>%
#               mutate(data_id = as.character(data_id))

ggplot(all.kUs, aes(x = rowname, y = value)) +
  geom_point(aes(color = name)) +
  geom_boxplot(outlier.shape = NULL) +
  coord_flip() + 
  theme_bw() +
  scale_y_continuous(limits = c(20, 56), breaks = seq(20, 56, by = 2),
                     minor_breaks = NULL) +
  scale_x_discrete(labels = NULL) + 
  labs(y = "kU (words/month)", x = "Subject")

```

This shows that for the majority of individuals, using 3, 4, or 5 points 
provide reasonable estimates, however, there are some individuals whose spread
is large.

These individuals don't seem to have a systematically low oldest visit or 
low largest inventory size.

However, it appears, as a rule of thumb, at least the final value should 
exceed $\approx100$, or
for simplicity's sake, $0.5 * W_i = 125$.

\newpage
# Lexical vs syntactic

Are the lexical vs. syntactic scores discriminable? (Subject selected at 
random from those with five visits.) Here, the curve is fit between 0 and 1, as
a proportion of items endorsed. This makes $W_i\approx{}0.37$.

This plot uses the 95% confidence interval rather than the SE.

```{r lexsyn, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3}

lexsyn <- read_csv("G:/My Drive/Research/MCDI/MCDI-analysis/data/other/lexsyn.csv")

three_or_more <- table(lexsyn$data_id)
ID3 <- names(three_or_more)[three_or_more >= 3]

lexsyn <- lexsyn %>%
  filter(data_id %in% ID3) %>%
  arrange(data_id, age) %>%
  select(-lex, -syn) %>%
  group_by(data_id, age) %>%
  pivot_longer(-c(data_id, age)) %>%
  filter(age > 0)

test <- filter(lexsyn, data_id == "116056")

# ggplot(test, aes(x = age, y = value, color = name)) +
#   geom_point() +
#   scale_x_continuous(limits = c(10, 40)) +
#   scale_y_continuous(limits = 0:1) +
#   geom_line()

gformula <- function(kg, W_0 = .Machine$double.eps, A = 1, adj = 0) {
  
  f <- function(.x) { W_0 * (A/W_0)^(1 - exp(-kg * (.x + adj) )) }
  return(f)
  
}

gminmax <- function(fit, min = TRUE) {
  
  kg <- coef(fit) %>% unname()
  se <- coef(summary(fit))[1, 2]
  ci <- 1.96 * se
  
  if (min)
    return(kg - ci)
  else
    return(kg + ci)
  
}

test.l <- gomp2.fit(filter(test, name == "lex.p"), response = "value", 
                    max = 1) 
test.s <- gomp2.fit(filter(test, name == "syn.p"), response = "value", 
                    max = 1) 

colors <- hue_pal()(2)

p.lxsy <- ggplot(NULL) +
            scale_x_continuous(limits = c(0, 60)) +
            geom_hline(yintercept = 1 / exp(1)) +
            labs(x = "Age (mo.)", y = "Proportion sub-score", 
                 color = "Subscore") +
            theme(legend.position = "below") + 
            scale_colour_discrete(labels = c("Lexical", "Syntax")) +
            geom_point(data = filter(test, name %in% c("lex.p", "syn.p")),
                       aes(x = age, y = value, color = name)) +
            stat_function(data = test,
                          fun = gformula(coef(test.l)), 
                          color = colors[1], size = 1) +
            stat_function(data = test,
                          fun = gformula(coef(test.s)), 
                          color = colors[2], size = 1) +
            geom_ribbon(data = predict.gomp(test.l, range = 10:60, max = 1,
                                                ci = TRUE),
                        aes(x = range, 
                            ymax = result.hi,
                            ymin = result.lo),
                        fill = colors[1], alpha = 0.25) +
            geom_ribbon(data = predict.gomp(test.s, range = 10:60, max = 1,
                                                ci = TRUE),
                        aes(x = range, 
                            ymax = result.hi,
                            ymin = result.lo),
                        fill = colors[2], alpha = 0.25) 

p.lxsy

png("plots/lexsyn-gompertz-presentation.png", width = 5.5, height = 5,
    res = 300, units = "in")

print(p.lxsy)

invisible( dev.off() )

```

The take-away here is that syntactic development occurs later, and even though
the SE is larger, the ranges do *not* overlap. It is the difference between
$T_i,lex$ and $T_i,syn$ that is interesting.

## Lexical vs. syntactic across subjects

Individuals in rows/columns (not one subject per row as above).

```{r LS plots, echo=FALSE, fig.height=3, fig.width=3, message=FALSE, warning=FALSE}

lexVsyn <- function(data) {
  
  lexical.fit <- gomp2.fit(filter(data, name == "lex.p"), 
                            response = "value", max = 1, max.iter = 500) 
  syntax.fit <- gomp2.fit(filter(data, name == "syn.p"), 
                          response = "value", max = 1, max.iter = 500) 
  
  colors <- hue_pal()(2)
  
  if (!is.na(lexical.fit) & !is.na(syntax.fit))
    plot <- ggplot(NULL) +
              scale_x_continuous(limits = c(0, 60)) +
              geom_hline(yintercept = 1 / exp(1)) +
              labs(x = "Age (mo.)", y = "Proportion sub-score", 
                   color = "Subscore") +
              theme(legend.position = "below") + 
              scale_colour_discrete(labels = c("Lexical", "Syntax")) +
              geom_point(data = filter(data, name != "SUM"),
                         aes(x = age, y = value, color = name)) +
              stat_function(data = test,
                            fun = gformula(coef(lexical.fit)), 
                            color = colors[1], size = 1) +
              stat_function(data = data,
                            fun = gformula(coef(syntax.fit)), 
                            color = colors[2], size = 1) +
              geom_ribbon(data = predict.gomp(lexical.fit, range = 10:60, max = 1,
                                                  ci = TRUE),
                          aes(x = range, 
                              ymax = result.hi,
                              ymin = result.lo),
                          fill = colors[1], alpha = 0.25) +
              geom_ribbon(data = predict.gomp(syntax.fit, range = 10:60, max = 1,
                                                  ci = TRUE),
                          aes(x = range, 
                              ymax = result.hi,
                              ymin = result.lo),
                          fill = colors[2], alpha = 0.25) +
              labs(title = unique(data$data_id))
  else
    plot <- NA
  
  return(plot)
  
}

for (i in ID3)
  print(lexVsyn(filter(lexsyn, data_id == i)))

```

## SE plots

```{r LS.vals, error=FALSE}

LS.vals <- data.frame(data_id = ID3) 
lex.fits <- lapply(LS.vals$data_id, 
                    function(x) gomp2.fit(filter(lexsyn,
                                                 data_id == x,
                                                 name == "lex.p"), 
                                          response = "value", max = 1, 
                                          max.iter = 500) )
syn.fits <- lapply(LS.vals$data_id, 
                    function(x) gomp2.fit(filter(lexsyn, 
                                                 data_id == x,
                                                 name == "syn.p"), 
                                          response = "value", max = 1, 
                                          max.iter = 500) )

# Get values
LS.vals <- LS.vals %>%
  mutate(
    lex = sapply(lex.fits, coef),
    lex.se = sapply(lex.fits, function(x) coef(summary(x))[1, 2]),
    syn = sapply(syn.fits, function(x) ifelse(class(x) != "nls", NA, coef(x))),
    syn.se = sapply(syn.fits, 
                    function(x) ifelse(class(x) != "nls", NA,
                                       coef(summary(x))[1, 2]))
  )

LS.vals %>%
  select(data_id, lex.m, syn.m) %>%
  pivot_longer(-data_id) %>%
  ggplot(aes(value, fill = name)) +
    geom_histogram(position = "dodge") +
    scale_y_continuous(breaks = 1:22) +
    theme_bw()

ggplot(LS.vals, aes(x = lex.m, y = syn.m)) +
  geom_point()

# Value at Wi 
W_i.1 <- 1 / exp(1)
LS.vals <- LS.vals %>%
  mutate(
    lex.m  = solve.gomp2(y = W_i.1, k_g = lex, A = 1),
    lex.lo = solve.gomp2(y = W_i.1, k_g = lex - (2 * lex.se), A = 1),
    lex.hi = solve.gomp2(y = W_i.1, k_g = lex + (2 * lex.se), A = 1),
    syn.m  = solve.gomp2(y = W_i.1, k_g = syn, A = 1),
    syn.lo = solve.gomp2(y = W_i.1, k_g = syn - (2 * syn.se), A = 1),
    syn_hi = solve.gomp2(y = W_i.1, k_g = syn + (2 * syn.se), A = 1),
    synMlex = syn.m - lex.m
  )

mean(LS.vals$synMlex, na.rm = TRUE)
sd(LS.vals$synMlex, na.rm = TRUE)

t.test(LS.vals$lex.m, LS.vals$syn.m, paired = TRUE)

ggplot()


```

```{r LS.vals_plot, message=FALSE, warning=FALSE, include=FALSE}

# LS.vals.melt <-  LS.vals %>%
#                   select(-syn.less.lex) %>%
#                   pivot_longer(-data_id)

png("plots/rangediff-presentation.png", width = 5, height = 5, units = "in",
    res = 300)

ggplot(LS.vals) +
  geom_pointrange(aes(x = data_id, 
                      y = 0, ymin = lex.hi - lex.m, ymax = lex.lo - lex.m), 
                  color = hue_pal()(2)[1], 
                  size = 0.75, fatten = 0.25) +
  geom_pointrange(aes(x = data_id, 
                      y = syn.m - lex.m, ymin = syn.hi - lex.m, 
                      ymax = syn.lo - lex.m), 
                  color = hue_pal()(2)[2], 
                  size = 0.75, fatten = 0.25) +
  coord_flip() +
  scale_x_discrete(labels = NULL) +
  scale_y_continuous(limits = c(-6, NA)) + 
  labs(y = "Time between {lex = Wi} and {syn = Wi} (mo)", x = "Participant")

dev.off()

png("plots/differences-presentation.png", width = 5, height = 5, units = "in",
    res = 300)

ggplot(LS.vals, aes(synMlex)) +
  geom_histogram(binwidth = .25) +
  labs(x = "Ti,syn - Ti,lex (mo.)") 


# Suppress dev.off() messages
invisible( dev.off() )

```


## All curves

```{r all-curves, echo=TRUE, message=FALSE, warning=FALSE}

LS.vals <- LS.vals %>%
            mutate(syn.m)


lexlines <- alply(as.matrix(LS.vals[, c("lex", "lex.m")]), 1, 
                  function(x) stat_function(fun = gformula(x[1], adj = x[2]), 
                                            color = "grey",
                                            alpha = .25) )

synlines <- alply(as.matrix(LS.vals[, c("syn", "lex.m")]), 1, 
                  function(x) stat_function(fun = gformula(x[1], adj = x[2]), 
                                            color = colors[2],
                                            alpha = 0.25) )

png("plots/lexsyn-lines.png", width = 5, height = 5, units = "in", res = 300)

ggplot(LS.vals) +
  scale_x_continuous(limits = c(-20, 40)) +
  scale_y_continuous(limits = c(0, 1)) + 
  geom_hline(yintercept = 1 / exp(1)) +
  labs(x = "Centered on T_i,lex", y = "Proportion sub-score") +
  lexlines + synlines

invisible( dev.off() )

```

\newpage
# Package info

```{r package-info}

version

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]

```

# References
