---
title: "Wordbank MB-CDI model improvement"
author: "Trevor K.M. Day"
output: 
  pdf_document:
    toc: true
---

# Setup

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	warning = FALSE
)

path <- "/Research/MCDI/MCDI-analysis/code/MNLFA"
locs <- c("G:/My Drive", "I:", "/Volumes")
for (i in locs)
  if (dir.exists(paste0(i, path)))
    setwd(paste0(i, path))

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lavaan))

source("../mcdi-setup.R")

wbws <- read_data("Wordbank/WS-scored.rds")$p 

show_model <- function(model){
  
  FMs <- fitMeasures(model, c("pvalue", "cfi", "tli", "rmsea", "srmr"))
  print(FMs)
  
  MIs <- modificationindices(model, standardized = TRUE, sort. = TRUE, 
                             minimum.value = 10) 
  
  print(nrow(MIs))
  print(head(MIs))
  
  return(FMs)
}

```

Packages loaded: `tidyverse` (1.3.1), `lavaan` (0.6-9)

# Data

Wordbank (WB) Words and Sentences (WS) data:

```{r data}
head(wbws) %>%
  mutate(
    across(where(is.numeric), ~round(.x, 3))
  )
```

# Initial models

The initial models are given below, with no cross-correlations.

```{r setup-formulae}

lex <- 'MR1 =~ animals + body_parts + clothing + food_drink + furniture_rooms + 
                games_routines + household + sounds + toys + vehicles'

syn <- 'MR2 =~ connecting_words + helping_verbs + pronouns + quantifiers + 
                question_words + COMPLEXITY'

# Fit measures to evaluate
fitms <- c("pvalue", "cfi", "tli", "rmsea", "srmr")

lex_models <- list()
lex_fitms <- list()
syn_models <- list()
syn_fitms <- list()

```

# Lexical modeling

## Step 1

```{r lex1}

lex_models[[1]] <- cfa(model = lex, data = wbws)
lex_fitms[[1]] <- show_model(lex_models[[1]])

```

This results in 37 MIs > 10, but only one < 1000.

## Step 2

```{r lex2}

lex[2] <- 'MR1 =~ animals + body_parts + clothing + food_drink + 
                  furniture_rooms + games_routines + household + sounds + 
                  toys + vehicles

            furniture_rooms ~~ household'

lex_models[[2]] <- cfa(model = lex[2], data = wbws)
lex_fitms[[2]] <- show_model(lex_models[[2]])

```

## Step 3

Now I'm going to add the four that exceed 200. 

```{r lex3}

lex[3] <- 'MR1 =~ animals + body_parts + clothing + food_drink + 
                  furniture_rooms + games_routines + household + sounds + 
                  toys + vehicles

            furniture_rooms ~~ household
            animals ~~ sounds
            animals ~~ vehicles
            games_routines ~~ sounds
            body_parts ~~ games_routines
'

lex_models[[3]] <- cfa(model = lex[3], data = wbws)
lex_fitms[[3]] <- show_model(lex_models[[3]])

```

## Model improvement

This removed four MIs from 35, resulting in only 23 that exceed 10. Let's take
a step back and evaluate model improvement.

```{r}

lex_fitms_all <- lex_fitms %>%
  unlist() %>%
  matrix(ncol = 3) %>%
  t() %>%
  as_tibble()

colnames(lex_fitms_all) <- fitms

lex_fitms_all <- lex_fitms_all %>%
  mutate(
    n = 1:nrow(lex_fitms_all)
  ) %>%
  pivot_longer(-n)

ggplot(lex_fitms_all, aes(x = n, y = value, color = name)) +
  geom_point() + 
  geom_line() + 
  geom_hline(yintercept = c(0.9, 0.08)) + 
  theme_bw()


```
CFI and TLI are both larger = better, and were above threshold to begin with.
SRMR and RMSEA are both smaller = better, and RMSEA seems to be trending the 
right way.  Chi-squared $p$-value is irrelevant at this sample size but is 
included for completeness. 

## Step 4

Let's try removing >200 again.


```{r lex4}

lex[4] <- 'MR1 =~ animals + body_parts + clothing + food_drink + 
                  furniture_rooms + games_routines + household + sounds + 
                  toys + vehicles

            furniture_rooms ~~ household
            animals ~~ sounds
            animals ~~ vehicles
            games_routines ~~ sounds
            body_parts ~~ games_routines
            sounds ~~ vehicles
            body_parts ~~ sounds
'

lex_models[[4]] <- cfa(model = lex[4], data = wbws)
lex_fitms[[4]] <- show_model(lex_models[[4]])

```

## Step 5

RMSEA is not quite sub-threshold, so now >100.

```{r lex5}

lex[5] <- 'MR1 =~ animals + body_parts + clothing + food_drink + 
                  furniture_rooms + games_routines + household + sounds + 
                  toys + vehicles

            furniture_rooms ~~ household
            animals ~~ sounds
            animals ~~ vehicles
            games_routines ~~ sounds
            body_parts ~~ games_routines
            sounds ~~ vehicles
            body_parts ~~ sounds
            body_parts ~~ furniture_rooms 
            furniture_rooms ~~ games_routines 
            clothing ~~ furniture_rooms
            body_parts ~~ household
'

lex_models[[5]] <- cfa(model = lex[5], data = wbws)
lex_fitms[[5]] <- show_model(lex_models[[5]])

lex_fitms_all <- lex_fitms %>%
  unlist() %>%
  matrix(ncol = length(lex_fitms)) %>%
  t() %>%
  as_tibble()

colnames(lex_fitms_all) <- fitms

lex_fitms_all <- lex_fitms_all %>%
  mutate(
    n = 1:nrow(lex_fitms_all)
  ) %>%
  pivot_longer(-n)

ggplot(lex_fitms_all, aes(x = n, y = value, color = name)) +
  geom_point() + 
  geom_line() + 
  geom_hline(yintercept = c(0.9, 0.08)) + 
  theme_bw()

```

This is the most parsimonious model that satisifes the criteria for CFI, TLI,
RMSEA, and SRMR. 

# Syntactic modeling

## Step 1 

```{r syn1}

syn_models[[1]] <- cfa(model = syn, data = wbws)
syn_fitms[[1]] <- show_model(syn_models[[1]])

```

This leaves only a single MI > 100 (to align with lexical).

## Step 2

```{r syn2}

syn[[2]] <- 'MR2 =~ connecting_words + helping_verbs + pronouns + quantifiers + 
                question_words + COMPLEXITY

              connecting_words ~~ helping_verbs
'

syn_models[[2]] <- cfa(model = syn[[2]], data = wbws)
syn_fitms[[2]] <- show_model(syn_models[[2]])

```

What luck, this works and satisifes the same criterion as lexical. I'm not going
to plot two points per measure. 

# Final models

```{r final-models}
writeLines(lex[length(lex)])

writeLines(syn[length(syn)])
```
```{r fit-stats}

final_model <- paste(lex[length(lex)], syn[length(syn)])
final_cfa <- cfa(final_model, wbws)

fitmeasures(final_cfa, fit.measures = fitms)

```

# Appendix: Session Info

```{r packages}
sessionInfo()
```
